---
title: "R Notebook"
output: html_notebook
---
Importing libraries
```{r}
library(ggplot2)
library(dplyr)
library(MCMCpack)
```

Read input and remove NA values

```{r}
data_all<-read.csv('winemag-data-130k-v2.csv')
data<-droplevels(na.omit(data_all))
```

Plotting *points* histogram
```{rcollapse=TRUE}
ggplot(data, aes(x=points)) + geom_histogram(aes(y=..density..),
                   binwidth=1,alpha=.5, position="dodge"
                   ) +
    geom_density(adjust = 5)
  
```
Filtering and grouping data
```{r}
data_cons <- data %>% 
  dplyr::filter(((country=='South Africa' & variety=='Sauvignon Blanc')|(country=='Chile' & variety=='Chardonnay')) & (price==15)) %>%
  dplyr::select(variety, points) %>%
  group_by(variety) %>% 
  dplyr::mutate(points.mean=mean(points), points.variance=var(points))
```
Comparing the two categories
```{r collapse=TRUE}
ggplot(data_cons) + geom_boxplot(aes(variety, points, fill = variety)) +
geom_jitter(aes(variety,points, shape = variety))
ggplot(data_cons, aes(color = variety, sample = rt(nrow(data_cons), df = 18)))  + geom_qq()
  
ggplot(data_cons, aes(x=points)) + geom_histogram(aes(y=..density..),
                   binwidth=0.5,alpha=.5, position="dodge"
                   )  +
  geom_vline(data=data_cons, aes(xintercept=points.mean),
               linetype="dashed", size=1) +
    geom_density(alpha=.2) +
  facet_grid(variety ~ .)  # Overlay with transparent density plot
```
Performing Welch T-test
```{r collapse=TRUE}
res <- t.test(points ~ variety, data=data_cons, alternative="less") # note absence of var.equal=TRUE
res
data_cons %>% group_by(variety) %>% summarise(n=n(), mean=mean(points), variance=var(points), min=min(points), max=max(points), sd=sd(points))
#stats <- data_cons %>% group_by(variety) %>% summarize(mean = mean(points), variance=var(points))
test <- function(){
  ntotal <- nrow(data_cons)
  nsamp <- 50 ## subset to sample
  ind_samp <- sample(1:ntotal, nsamp, replace = TRUE) ## which datapoints to sample
  df_samp <- data_cons[ind_samp, ] ## which datapoints to sample
  df_samp <- droplevels(df_samp)
  return(t.test(points ~ variety, data=df_samp))
}
test_samples <- replicate(500,test()$statistic)
#range(test_samples)[1]
pts = seq(-8,8,length=500)
x <- plot(density(test_samples),col='red',type='l', main = "Welch T density")
abline(v = res$statistic)
text(c(-2, res$statistic-0.15), c(0.2, 0.15), labels = c("", res$statistic), srt=90)
#lines(density(test_samples))
#qqplot(test_samples,rt(1000,df=18))
#abline(0,1)
```
Define prior values and gibbs function
```{r collapse=TRUE}
qnorm(c(0.025, 0.975), 85, 10)
compare_2_gibbs <- function(y, ind, mu0 = 85, tau0 = 1/49, del0 = 1, gamma0 = 1/16, a0 = 50, b0 = 150, maxiter = 5000) {
  y1 <- y[ind == 'Sauvignon Blanc']
  y2 <- y[ind == 'Chardonnay']
  n1 <- length(y1)
  n2 <- length(y2)
  ##### starting values
  mu <- (mean(y1) + mean(y2)) / 2
  del <- (mean(y1) - mean(y2)) / 2
  mat_store <- matrix(0, nrow = maxiter, ncol = 3)
  #####
  ##### Gibbs sampler
  an <- a0 + (n1 + n2)/2
  for(s in 1 : maxiter) {
    ##update tau
    bn <- b0 + 0.5 * (sum((y1 - mu - del) ^ 2) + sum((y2 - mu + del) ^ 2))
    tau <- rgamma(1, an, bn)
    ##
    ##update mu
    taun <- tau0 + tau * (n1 + n2)
    mun <- (tau0 * mu0 + tau * (sum(y1 - del) + sum(y2 + del))) / taun
    mu <- rnorm(1, mun, sqrt(1/taun))
    ##
    ##update del
    gamman <- gamma0 + tau*(n1 + n2)
    deln <- ( del0 * tau0 + tau * (sum(y1 - mu) - sum(y2 - mu))) / gamman
    del<-rnorm(1, deln, sqrt(1/gamman))
    ##
    ## store parameter values
    mat_store[s, ] <- c(mu, del, tau)
  }
  colnames(mat_store) <- c("mu", "del", "tau")
  return(mat_store)
}
```
Perform Gibbs sampling on data. Plot MCMC stats.


```{r}
fit <- compare_2_gibbs(data_cons$points, as.factor(data_cons$variety))
plot(as.mcmc(fit))
acf(fit)
raftery.diag(fit)

```
Some more plots...
```{r collapse=TRUE}
par(mfrow = c(2, 2))
plot(density(fit[, 1]), main = "Posterior for mean of difference",
xlab = expression(mu))
plot(density(fit[, 2]), main = "Posterior for precision of difference",
xlab = expression(tau))
plot(density(1/sqrt(fit[, 2]), na.rm=T), main = "Posterior for sd of difference",
xlab = expression(sigma))
plot(fit[, 1], 1/sqrt(fit[, 2]),
main = "Joint posterior for mean and sd of difference",
xlab = expression(mu), ylab = expression(sigma))
```
Numeric values for mu, del and tau
```{r collapse=TRUE}
apply(fit, 2, mean)
apply(fit, 2, sd)
mean(1/sqrt(fit[, 3]))
sd(1/sqrt(fit[, 3]))
```
Generate random samples and plot difference histogram
```{r collapse=TRUE}
y1_sim <- rnorm(5000, fit[, 1] + fit[, 2], sd = 1/sqrt(fit[, 3]))
y2_sim <- rnorm(5000, fit[, 1] - fit[, 2], sd = 1/sqrt(fit[, 3]))
ggplot(data.frame(y_sim_diff = y1_sim - y2_sim)) + stat_bin(aes(y_sim_diff), binwidth = 1, alpha = 0.7)
```
Calculate probability of difference of mean
```{r}
mean(y1_sim > y2_sim)
ggplot(data.frame(y1_sim, y2_sim)) + geom_point(aes(y1_sim, y2_sim), alpha = 0.3) +
geom_abline(slope = 1, intercept = 0)
```
```{r}
a0 <- 1
b0 <- 100
tau_range <- qgamma( c(0.005, 0.995), a0, rate = b0)
## N.B., you can define dgamma in terms of scale or rate parameter
## make sure you specify correctly!
xseq <- seq(tau_range[1], tau_range[2], length.out = 100)
y <- dgamma(xseq, a0, b0)
ggplot(data.frame(xseq, y)) + geom_line(aes(xseq, y)) #+ geom_vline(xintercept = 0.0688)
#xlab = expression(tau), main = "Prior for precision parameter")
#abline(v = 0.01, col = 2, lty = 2, lwd = 2) ## true value of tau'


mu <- 5
a_n <- 45.5
b_n <- 205

tau_range <- qgamma( c(0.005, 0.995), a_n, rate = b_n)
xseq <- seq(tau_range[1], tau_range[2], length.out = 1000)
plot(xseq, dgamma(xseq, a_n, rate = b_n), type="l", xlab = expression(tau),
main = "Posterior for precision parameter")
abline(v = 0.01, col = 2, lty = 2, lwd = 2) ## true value of tau

## prior density plot for standard deviation sigma
tau_post_samp <- rgamma(1000, a_n, rate = b_n)
hist(1/sqrt(tau_post_samp), main = "Posterior for standard deviation",
xlab = expression(sigma), freq = FALSE)
lines(density(1/sqrt(tau_post_samp)))
abline(v = 10, col = 2, lty = 2, lwd = 2) ## true value of sd
```

